---
title: "Top100 School Districts"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)
library(janitor)
library(plotly)
library(forcats)
library(DT)
library(jsonlite)
options(scipen = 999)
df <- data_frame(state.abb, state.name)
source("nces.R")
source("nyc_school_districts.R")
```

# School districts from Acfrs database 2020, 2021

```{r echo = FALSE}
db2020 <- readRDS("data/data_from_dbsite_2020.RDS") 
db2021 <- readRDS("data/data_from_dbsite_2021.RDS")
db2022 <- readRDS("data/data_from_dbsite_2022.RDS")

sd_20_21_22 <- db2020 %>% 
  rbind(db2021) %>% 
  rbind(db2022) %>% 
  filter(category == "School District") %>% 
  mutate(id = as.character(id)) %>% 
  select(state, name,id, year, 
         total_liabilities, net_pension_liability, net_opeb_liability, 
         net_pension_assets, net_opeb_assets,
         expenses, revenues)
```

## Top School Districts in ACFRs

```{r echo = FALSE}
# Dictionary created in this project, file "dictionary_name_ncesID_acfrsID.qmd"
#TODO: 
# those in list 101-200 nces but NOT in acfrs yet

####
dictionary <- readRDS("dictionary.RDS")
sub_dictionary <- data.frame(
                  ncesID = c("4703180"), #ncesID == "4703180" ~ "107203", # Davidson County Board of Education, previously id = 1267426
                  name = c("Metropolitan Nashville Public Schools"),
                  id = c("107203"),
                  state = c("TN")
)

dict_top100 <- dictionary %>% 
  rbind(sub_dictionary) %>% 
  filter(ncesID %in% top_sd_nces$ncesID) %>% 
  mutate(id = case_when(ncesID == "5101260" ~ "1267421", # why previous id = "1250807", # Fairfax County Public Schools
                   ncesID == "5102250" ~ "1250804", #Loudoun County Public Schools	
                   ncesID == "5103130"~	"1250808",# Prince William County Public Schools		VA
                   ncesID == "0803360" ~ "1237862", #City and County of Denver School District No. 1
                   ncesID == "0102370" ~ "1017229",#Mobile County Board of School Commissioners	
                   ncesID == "5103840" ~ "1265776", #School Board of the City of Virginia Beach	
                   ncesID == "5100840" ~ "1267422", #VA	Chesterfield County Public Schools	1267422
                   ncesID == "5101890" ~ "1267423", #VA	Henrico County Public Schools	1267423
                   ncesID == "2502790" ~ "1267424", #MA	Boston Public Schools	
                   ncesID == "3174820" ~ "35480", #NE Omaha City School District 1 = 	Douglas County School District No. 0001	35480
                   ncesID == "2601103" ~ "161679", # MI Detroit Public Schools
                   ncesID == "1602100" ~ "32700", #ID	JOINT SCHOOL DISTRICT NO. 2
         TRUE ~ as.character(id))) 

# remain 5 cases of NYC not yet in dictionary: 3600103" "3600077" "3600098" "3600087" "3600151"
# these will be treated separately below
setdiff(top_sd_nces$ncesID, dict_top100$ncesID)
top_sd_nces %>% filter(ncesID %in% c("3600103", "3600077", "3600098", "3600087", "3600151"))
```


```{r}
dict_top101_200 <- dictionary %>% 
  rbind(sub_dictionary) %>% 
  filter(ncesID %in% top_sd_nces_102_200$ncesID)

#in top nces but not in dictionary: NYC 27, 28, 30, 11, 25, 21, 22, 9 are handled separately below

anti_join(top_sd_nces_102_200, dict_top101_200,  by= "ncesID")

# need to check: 
#2733840	ST. PAUL PUBLIC SCHOOL DISTRICT	MN
#4703030	Montgomery County	TN
#1713710	SD U-46	IL
#0802340	Aurora Joint District No. 28 of the counties of Adams and A	CO
#4110040	Portland SD 1J	OR
```

```{r}
top100_sd_3years <- sd_20_21_22 %>% 
  filter(id %in% dict_top100$id) %>% select(-name) %>% 
  left_join(dict_top100, by = c("id", "state"))  %>% 
  
  #bind with NYC
  rbind(nyc_top5_20_21_22) %>% 
  rename(state.abb = state) %>% 
  left_join(df) %>% 
  
  select(state.abb, state.name, ncesID, id, year, name, 
         total_liabilities, net_pension_liability, net_opeb_liability, 
         net_pension_assets, net_opeb_assets,
         expenses, revenues) %>% drop_na(ncesID) %>% 
  #join with nces to get county, city info
  left_join(nces, by = "ncesID") %>% select(-c(state)) 
#%>% write.csv("output/top100_sd_3years_RON.csv")
 
#TODO: check back missing: 
#GA Clayton County Board of education.
#https://www.clayton.k12.ga.us/departments/business-services/financial-reports

top100_sd_3years %>% 
  add_count(ncesID) %>% filter(n < 3) %>% 
  select(state.abb, ncesID, year, name, n)

top_sd_nces %>% filter(ncesID %in% setdiff(top_sd_nces$ncesID, top100_sd_3years$ncesID))

```


```{r}
#saveRDS(top100_sd_3years, "top100_sd_3years.RDS")
exportJSON <- toJSON(top100_sd_3years)
write(exportJSON, "output/top100_sd_3years.json")

# <- fromJSON("top100_sd_3years.json")
top100_sd_3years %>% write.csv("output/top100_sd_3years.csv")

#TODO: ask Geoff about this revenues = expense cases
top100_sd_3years %>% arrange(desc(students)) %>% 
  filter(expenses == revenues) 

#
top100_sd_3years %>% filter(name == "Prince George’S County Public Schools")
```

# sd101_200_3years


```{r}
top_sd_nces_102_200
sd101_200_3years <- sd_20_21_22 %>% 
  filter(id %in% dict_top101_200$id) %>% select(-name) %>% 
  left_join(dict_top101_200, by = c("id", "state")) %>% 
  rename(state.abb = state) %>% 

  left_join(df) %>% 
  #join with nces to get county, city info
  left_join(nces, by = "ncesID") %>% select(-c(state))
 
exportJSON <- toJSON(sd101_200_3years)
write(exportJSON, "output/sd101_200_3years.json")

# <- fromJSON("top100_sd_3years.json")
sd101_200_3years %>% write.csv("output/sd101_200_3years.csv")

# check missing
sd101_200_3years %>% add_count(ncesID) %>% select(state.abb, name, year, n) %>% 
  filter(n ==2)
```

 
```{r}
data.frame(
all <- sd_20_21_22 %>% mutate(all_NPL = sum(total_liabilities)) %>% select(all_NPL) %>% distinct(),

top100 <- top100_sd_3years %>% mutate(all_NPL = sum(total_liabilities)) %>% select(all_NPL) %>% distinct(),

next100 <- sd101_200_3years %>% mutate(all_NPL = sum(total_liabilities)) %>% select(all_NPL) %>% distinct()
) %>% format(big.mark = ",")

data.frame(all, top100, next100)
```


## Delaware
Delaware has school district info in the state ACFR: FY 2021 ACFR (delaware.gov). Balance sheet info is on p. 166-167 and activities are on p. 168-169.
--> manually inserted these numbers to excel file. 
--> Where to find net pension, net OPEB for K12? (State of Delaware in page 20)
--> Don't do this yet. Let me figure out why school districts are only listed in the statistical section--I'm not sure whether the liabilities are consolidated with the state's.

## Davidson County TN & Portland SD 1J OR

```{r warning=FALSE, message=FALSE}
sd_96_2020 <- top_sd_nces %>% 
  left_join(top_sd_acfrs) %>% # to get id
  left_join(sd2020_db, by = "id") %>% # need to specify join by=id, if not, some rows in sd2020_db loose values (eg. id == 196828) 
  
  select(-c(name.y, name.x, state.y)) %>% 
  rename(state = state.x) %>% 

  # take out sd whose acfrs are not known to date
  filter(!nces_original_name %in% c("Boston", "CHESTERFIELD CO PBLC SCHS", "HENRICO CO PBLC SCHS", "OMAHA PUBLIC SCHOOLS")) %>% 
  
  # take out 5 sd of NYC to later rbind with nyc_20_21_5sd
  filter(state != "NY") %>% 
  
  # Collect data for some sd manually 
  mutate(year = 2020,
  total_liabilities = case_when(nces_original_name == "Davidson County"& state == "TN" ~ 0,
                                nces_original_name == "Portland SD 1J"& state == "OR" ~ 1882236000, 
                                TRUE ~ total_liabilities), #page 34
  
  net_pension_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                                    nces_original_name == "Portland SD 1J"& state == "OR" ~ 306140000, 
                                    TRUE ~ net_pension_liability), # page 34, Net pension liability-PERS
  
  net_opeb_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                                 nces_original_name == "Portland SD 1J"& state == "OR" ~ 81319999,
                                 TRUE ~ net_opeb_liability), # page 34, TOTAL OPEB liability-RHIS
  
  expenses = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ 824384000, 
                       TRUE ~ expenses),#page 35
                       
  
  revenues = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ (8271000 + 90072000 + 11368000 + 823400000), #page 35 (charges + operating + grant + Total general revenues)
                       TRUE ~ revenues))
```

```{r}
sd_96_2021 <- top_sd_nces %>% 
  left_join(top_sd_acfrs) %>% # to get id
  left_join(sd2021_db, by = "id") %>% # NOTE: filter(is.na(total_liabilities)) -> some have id carried from 2020, but no values for 2021
  select(-c(name.y, name.x, state.y)) %>% 
  rename(state = state.x) %>%  
  
  # take out sd whose acfrs are not known to date
  filter(!nces_original_name %in% c("Boston", "CHESTERFIELD CO PBLC SCHS", "HENRICO CO PBLC SCHS", "OMAHA PUBLIC SCHOOLS")) %>% 
  
  # take out 5 sd of NYC to later rbind with nyc_20_21_5sd
  filter(state != "NY") %>% 
  
  mutate(year = 2021,
  total_liabilities = case_when(nces_original_name == "Davidson County"& state == "TN" ~ 0,
                                nces_original_name == "Portland SD 1J"& state == "OR" ~ 2380580000,
                                TRUE ~ total_liabilities),
  
  net_pension_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 84041613,
                                    nces_original_name == "Portland SD 1J"& state == "OR" ~ 521329000,
                                    TRUE ~ net_pension_liability), # page 32, Net pension liability-PERS
  
  net_opeb_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 1061732808, # page B-93, column School professional employee
                                 nces_original_name == "Portland SD 1J"& state == "OR" ~ 78581000,
                                 TRUE ~ net_opeb_liability), # page 32, TOTAL OPEB liability-RHIS
  
  expenses = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 1195361854,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ 931870000,
                       TRUE ~ expenses), #page 33
  
  revenues = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 488401680,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ (1914000 + 113835000 + 1319000 + 822868000), # Geoff: need to take sum 4 of these: charges, operating, capital grant and general revenues)
                       TRUE ~ revenues) #page 33
)

```




