---
title: "Top100 School Districts"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(stringr)
library(tidyr)
library(janitor)
library(forcats)
source("nces.R")
source("general_purpose.R")
source("nyc_school_districts.R")
options(scipen = 999)
df <- data_frame(state.abb, state.name)
```

# School districts from Acfrs database 2020, 2021

```{r echo = FALSE}
school_districts <- readRDS("data/acfrs_data.RDS") %>% 
  filter(category == "School District") %>% 
  mutate(id = as.character(id)) %>% 
  select(state, any_of(fields_to_select))
```

## Dictionary ACFRs ID ~ NCES ID

```{r echo = FALSE}
dictionary <- readRDS("dictionary.RDS") %>% 
  add_row(ncesID = c("4703180"), #ncesID == "4703180" ~ "107203", # Davidson County Board of Education, previously id = 1267426
                  name = c("Metropolitan Nashville Public Schools"),
                  id = c("107203"),
                  state = c("TN")) %>% 
  
#fixing id (old dictionary reflects old acfrs id --> update these new acfrs id currently showing on database)
  mutate(id = case_when(ncesID == "5101260" ~ "1267421", # why previous id = "1250807", # Fairfax County Public Schools
                   ncesID == "5102250" ~ "1250804", #Loudoun County Public Schools	
                   ncesID == "5103130"~	"1250808",# Prince William County Public Schools		VA
                   ncesID == "0803360" ~ "1237862", #City and County of Denver School District No. 1
                   ncesID == "0102370" ~ "1017229",#Mobile County Board of School Commissioners	
                   ncesID == "5103840" ~ "1265776", #School Board of the City of Virginia Beach	
                   ncesID == "5100840" ~ "1267422", #VA	Chesterfield County Public Schools	1267422
                   ncesID == "5101890" ~ "1267423", #VA	Henrico County Public Schools	1267423
                   ncesID == "2502790" ~ "1267424", #MA	Boston Public Schools	
                   ncesID == "3174820" ~ "35480", #NE Omaha City School District 1 = 	Douglas County School District No. 0001	35480
                   ncesID == "2601103" ~ "161679", # MI Detroit Public Schools
                   ncesID == "1602100" ~ "32700", #ID	JOINT SCHOOL DISTRICT NO. 2
         TRUE ~ as.character(id))) 

```


```{r}
nces %>% select(state.abb, ncesID, enrollment_20, enrollment_21, enrollment_22) -> nces_enrollment

school_districts %>% select(state, name, id) %>% distinct() %>% 
  left_join(dictionary) %>% 
  left_join(nces_enrollment) %>% 
  filter(!is.na(ncesID)) %>% 
  mutate(government_level = "school district") %>% 
  write.csv("output/school district_enrollment_20_21_22.csv")
```

# Top 100 SD
```{r}
# filter only top 100
dict_top100_ELSI <- dictionary %>% 
  filter(ncesID %in% top_schools_by_year$ncesID) %>% 
  drop_na(id)

top100_school_districts <- school_districts %>% 
  filter(id %in% dict_top100_ELSI$id) %>% select(-name) %>% 
  left_join(dict_top100_ELSI, by = c("id", "state")) %>% 
  rename(state.abb = state) %>% 
  
  #bind with NYC
  rbind(nyc_top5_20_21_22) %>% 
  left_join(df) %>% 

  #join with nces to get county, city info
  left_join(nces, by = c("ncesID", "state.abb", "state.name")) %>% 
  rename(name = name.x) %>% 
  select(-name.y)

# Note: new york city geographic district #10 is not in the list of top 100 by enrollment 2022
top_schools_by_year %>% 
  mutate(name = str_to_lower(name)) %>% 
  filter(str_detect(name, "new york"))
 
#TODO: check back missing: 
#GA Clayton County Board of education.
#https://www.clayton.k12.ga.us/departments/business-services/financial-reports

top100_school_districts %>% 
  add_count(ncesID) %>% filter(n < 3) %>% 
  select(state.abb, ncesID, year, name, n)


```

# Export result
```{r}
saveRDS(top100_school_districts, "top100_sd.RDS")

top100_school_districts %>% write.csv("output/top100_sd.csv")
school_districts %>% write.csv("output/all_schooldistricts_3years.csv")
#TODO: ask Geoff about this revenues = expense cases
top100_school_districts %>% 
  filter(expenses == revenues) 


school_districts
```

## Delaware
Delaware has school district info in the state ACFR: FY 2021 ACFR (delaware.gov). Balance sheet info is on p. 166-167 and activities are on p. 168-169.
--> manually inserted these numbers to excel file. 
--> Where to find net pension, net OPEB for K12? (State of Delaware in page 20)
--> Don't do this yet. Let me figure out why school districts are only listed in the statistical section--I'm not sure whether the liabilities are consolidated with the state's.

## Davidson County TN & Portland SD 1J OR

```{r warning=FALSE, message=FALSE}
sd_96_2020 <- top_sd_nces %>% 
  left_join(top_sd_acfrs) %>% # to get id
  left_join(sd2020_db, by = "id") %>% # need to specify join by=id, if not, some rows in sd2020_db loose values (eg. id == 196828) 
  
  select(-c(name.y, name.x, state.y)) %>% 
  rename(state = state.x) %>% 

  # take out sd whose acfrs are not known to date
  filter(!nces_original_name %in% c("Boston", "CHESTERFIELD CO PBLC SCHS", "HENRICO CO PBLC SCHS", "OMAHA PUBLIC SCHOOLS")) %>% 
  
  # take out 5 sd of NYC to later rbind with nyc_20_21_5sd
  filter(state != "NY") %>% 
  
  # Collect data for some sd manually 
  mutate(year = 2020,
  total_liabilities = case_when(nces_original_name == "Davidson County"& state == "TN" ~ 0,
                                nces_original_name == "Portland SD 1J"& state == "OR" ~ 1882236000, 
                                TRUE ~ total_liabilities), #page 34
  
  net_pension_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                                    nces_original_name == "Portland SD 1J"& state == "OR" ~ 306140000, 
                                    TRUE ~ net_pension_liability), # page 34, Net pension liability-PERS
  
  net_opeb_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                                 nces_original_name == "Portland SD 1J"& state == "OR" ~ 81319999,
                                 TRUE ~ net_opeb_liability), # page 34, TOTAL OPEB liability-RHIS
  
  expenses = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ 824384000, 
                       TRUE ~ expenses),#page 35
                       
  
  revenues = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 0,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ (8271000 + 90072000 + 11368000 + 823400000), #page 35 (charges + operating + grant + Total general revenues)
                       TRUE ~ revenues))
```

```{r}
sd_96_2021 <- top_sd_nces %>% 
  left_join(top_sd_acfrs) %>% # to get id
  left_join(sd2021_db, by = "id") %>% # NOTE: filter(is.na(total_liabilities)) -> some have id carried from 2020, but no values for 2021
  select(-c(name.y, name.x, state.y)) %>% 
  rename(state = state.x) %>%  
  
  # take out sd whose acfrs are not known to date
  filter(!nces_original_name %in% c("Boston", "CHESTERFIELD CO PBLC SCHS", "HENRICO CO PBLC SCHS", "OMAHA PUBLIC SCHOOLS")) %>% 
  
  # take out 5 sd of NYC to later rbind with nyc_20_21_5sd
  filter(state != "NY") %>% 
  
  mutate(year = 2021,
  total_liabilities = case_when(nces_original_name == "Davidson County"& state == "TN" ~ 0,
                                nces_original_name == "Portland SD 1J"& state == "OR" ~ 2380580000,
                                TRUE ~ total_liabilities),
  
  net_pension_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 84041613,
                                    nces_original_name == "Portland SD 1J"& state == "OR" ~ 521329000,
                                    TRUE ~ net_pension_liability), # page 32, Net pension liability-PERS
  
  net_opeb_liability = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 1061732808, # page B-93, column School professional employee
                                 nces_original_name == "Portland SD 1J"& state == "OR" ~ 78581000,
                                 TRUE ~ net_opeb_liability), # page 32, TOTAL OPEB liability-RHIS
  
  expenses = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 1195361854,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ 931870000,
                       TRUE ~ expenses), #page 33
  
  revenues = case_when(nces_original_name == "Davidson County" & state == "TN" ~ 488401680,
                       nces_original_name == "Portland SD 1J"& state == "OR" ~ (1914000 + 113835000 + 1319000 + 822868000), # Geoff: need to take sum 4 of these: charges, operating, capital grant and general revenues)
                       TRUE ~ revenues) #page 33
)

```




