---
title: "testing_validation_analysis"
format: html
editor: visual
---


```{r}
source("data_processing/testing_validation.R")
```
# Test 1
## Counties

Hgarb did not determine total liabilities for JEFFERSON COUNTY 2023:  p11/61
report on cash basis, does not follow GAAP, no accrued liabilities
```{r}
test1_county %>% filter(year == 2023)

```

View all those failed test 1 in previous year
```{r}
test1_county %>% 
  arrange(desc(population)) 
  View()
```

## Municipalities

City of Lawrenceburg, Tennessee: 
Net Opeb, Net pension, total liabilites are all correct. Bonds = 69,370,682 taken from Changes in long term liabilities, summing up from pages 53,54,55.

Should only sum up these 2. Electric and water are already included. 
26,972,749 + 18,733,918 

```{r}
test1_municipality
```

AR jonesboro 2021, net pension on page 25= 19631568
total liabilities = current liability = 2,021,274 (sum of general, street, other funds) --> acceptable to calculate this way. 
Not follow GAAP. 


## School
```{r}
test1_school_districts %>% filter(total_liabilities != 0)
```
clover park school district no. 400
p62: should be 127,906,640.77 + current lia
Long term liabilities reported separately from balance. 


AR Pulsaki 2023
Calculation of total liabilities = 154977316 on pa20 + current 22642907


# Test 2

```{r}
View(test2_state_all)
```

```{r}
View(test2_county_all)
```

CRANE COUNTY, TEXAS: total_liabilities/total_assets > 200
p 70/93: The County’s net pension liability / (asset) of $114,769 --> wrong. Where to find total assets?

--> Don't report any assets

```{r}
test2_municipality_all %>% filter(total_assets != 0) %>% 
  arrange(desc(pct)) %>% View()

```

City of Dysart: Not following GAAP, 3929 is only assets we know about. 
Do not provide a balance sheet. 
```{r}
test2_school_districts %>% arrange(desc(pct)) %>% View()
```

# Test 3
## State
```{r}
test3_state_all
```

```{r}
View(test3_county_all)


```


```{r}

View(test3_municipality_all)


```


```{r}

View(test3_school_districts_all)
```

